---
title: "Activities of Daily Living (ADLs)"
author: "Tilen Rupar"
date: "23 6 2021"
output: pdf_document
---

```{r, echo=F, results='hide'}
setwd("E:/R/ADL/UCI ADL Binary Dataset")
Sys.setlocale("LC_TIME", "C")
```

```{r setup, warning = F, message=F}
library(tidyverse)
library(readr)
library(lubridate)
library(caret)
library(ggplot2)
library(class)
```

# Project goals
I will try to predict human activity based on sensory data with machine learning algorithm.

The machine learning algorithm is a supervised time series classification model. It uses POSIXct class (timestamp) to classify human activity, based on the time of the day. 

Data: [UCI Machine learning repository: ADL data](https://archive.ics.uci.edu/ml/datasets/Activities+of+Daily+Living+(ADLs)+Recognition+Using+Binary+Sensors)

```{r, echo=F}
na.omit(scan("OrdonezA_Description.txt", 
             sep = "\t", 
             what = "raw", 
             na.strings = "", 
             blank.lines.skip = T, ))

```


```{r Import data, warning = F, message=F}
#ADL.uci_file <- download.file('https://archive.ics.uci.edu/ml/machine-learning-databases/00271/UCI ADL Binary Dataset.zip', destfile = "./ADL.zip")
#unzip("./ADL.zip")

# Activities of Daily Living (ADLs) Recognition Using Binary Sensors Data Set (User "A")
ADL <- read.table("OrdonezA_ADLs.txt", skip = 2)
colnames(ADL) <- read.table("OrdonezA_ADLs.txt", nrows = 1)

#Creating a date time object for merging with the sensor dataset 
ADL$Start_Date_time <- as.POSIXct(paste(ADL$Start, ADL$time), format="%Y-%m-%d %H:%M:%S")
ADL <- subset(ADL, select = c(Activity, Start_Date_time))

#Reading the sensor dataset
ADL_sensor <- read.table("OrdonezA_Sensors.txt", skip = 2)
colnames(ADL_sensor) <- read.table("OrdonezA_Sensors.txt", nrows = 1)
ADL_sensor$Start_Date_time<- as.POSIXct(paste(ADL_sensor$Start, ADL_sensor$time), 
                                        format="%Y-%m-%d %H:%M:%S")
ADL_sensor <- subset(ADL_sensor, select = c(Start_Date_time, Location, Type, Place))
# Merging the Activity and sensor data
ADL <- merge(x = ADL, y = ADL_sensor, by = intersect(names(ADL), names(ADL_sensor)))
#add a day name, 
ADL$Day <- as.factor(weekdays(as.Date(ADL$Start)))
# a Dummy for weekend 
ADL$isWKND <- as.double(ADL$Day %in% c("Saturday", "Sunday"))

# and Time_of_day labels
# create breaks
breaks <- hour(hm("00:00", "6:00", "12:00", "18:00", "23:59"))
# labels for the breaks
labels <- c("Night", "Morning", "Afternoon", "Evening")
ADL$Time_of_day <- cut(x=hour(ADL$Start_Date_time), breaks = breaks, labels = labels, include.lowest=TRUE)
```

## Imported data:
```{r}
head(ADL)
```

We are interested in predicting Activity, based on Start and End time. Thus we turn the character vector "Activity" in factor with 9 levels: 
```{r, echo= FALSE}
unique(ADL$Activity)
``` 

# We visualize the Activity data: 
```{r}
ggplot(ADL, aes(x=Activity)) +
  geom_bar()
```

For the modelling purposes we need to convert the variables into numeric objects. 

```{r Structuring the data}
ADL_modData <- data.frame(Activity = as.factor(ADL$Activity),
                          Location = as.numeric(as.factor(ADL$Location)),
                          Type = as.numeric(as.factor(ADL$Type)),
                          Place = as.numeric(as.factor(ADL$Place)),
                          Day = as.numeric(ADL$Day),
                          Time_of_day = as.numeric(ADL$Time_of_day),
                          isWKND = ADL$isWKND)
str(ADL_modData)
```

Data partitioning (using caret package) - 0.8 data goes into training

```{r}
set.seed(100)
sample_rows <- as.vector(createDataPartition(ADL_modData$Activity, p=0.8, list = F))
ADL_train <- ADL_modData[sample_rows,]
ADL_test <- ADL_modData[-sample_rows,]
```

Applying the **k-Nearest neighbours model** to the ADL data 

```{r}
ADL_kNN <- knn(ADL_train[,-1], ADL_test[,-1], cl=ADL_train$Activity, k = 7)
```

# Creating the confusion matrix for the kNN model (Accuracy check)

```{r}
table(ADL_test$Activity, ADL_kNN)
```

## Next, we check the accuracy rate of the kNN model:

```{r}
mean(ADL_test$Activity == ADL_kNN)
```

The kNN model predicted with **86,4% accuracy rate**.

Next, we try to fit the **random forest model** to the data.

```{r}
#Before applying machine learning algorithms to train our model, let us first tune the cross-validation parameters. 
#We expect our out-of-sample error to be low because 5-fold CV should take its effect and avoid overfitting.
fitControl <- trainControl(method = "cv", number = 5, returnResamp = "all")

#Random forest model
ADL_rf <- train(Activity ~ Time_of_day + isWKND + Day + Location + Type + Place, method="rf", data=ADL_train,ntree=1000, tuneGrid=data.frame(.mtry = 3))
```

# Creating a prediction based on RF training data and analyzing the confusion matrix

```{r}
predictedADL_rf <- predict(ADL_rf, ADL_test)

confusionMatrix(ADL_test$Activity, predictedADL_rf)
```

AS we can see, the random forest model did not capture the data successfully,  the accuracy is 94,59%. 

# Presentation of the predicted vs. actual Activity 

```{r}

data.frame(Actual_data = ADL_test$Activity, Random_Forest_prediction = predictedADL_rf, kNN_prediction = ADL_kNN)
```

